---
title: "DiViMe tools with `avutils`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
  # pdf_document
vignette: >
  %\VignetteIndexEntry{divime}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)
```

```{r setup, message=FALSE}
library(avutils)
```

# Overview 

The idea here is to provide R functions that wrap up the shell commands normally used to invoke the virtual machine with the DiViMe tools. The functions are by and large calls using the `system2()` functions (or `system()`), with some processing around these calls. There is now also an early version of a shiny app (try `divime_app()`).

```{r, echo=FALSE, fig.width=8, fig.height=4.5}
par(family = "serif", mar = c(0, 0, 0, 0))
plot(0, 0, type = "n", xlim = c(0, 50), ylim = c(20, 0), axes = FALSE, ann = FALSE)
textcex <- 0.4
r <- FALSE

cds <- c(1, 4, 9, 2)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "speech activity detection", cex = textcex)

cds <- c(1, 6, 9, 8)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "divime_sad()", cex = textcex, family = "mono")

cds <- c(1, 10, 9, 14)
labs <- "noisemesSad_X.rttm\nopensmileSad_X.rttm\ntocomboSad_X.rttm"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)

cds <- c(1, 16, 9, 20)
labs <- "noisemesSad.sh\nopensmileSad.sh\ntocomboSad.sh"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)



# diarization -------------------------------------------------------------

cds <- c(11, 4, 19, 2)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "talker diarization", cex = textcex)

cds <- c(11, 6, 19, 8)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "divime_diarization(speech_annos = ...)", cex = textcex, family = "mono")

cds <- c(11, 10, 19, 14)
labs <- "diartk_noisemesSad_X.rttm\ndiartk_opensmileSad_X.rttm\ndiartk_tocomboSad_X.rttm"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)

cds <- c(11, 16, 19, 20)
labs <- "diartk.sh ... noisemesSad\ndiartk.sh ... opensmileSad\ndiartk.sh ... tocomboSad"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)

# talker type -------------------------------------------------------------

cds <- c(21, 4, 29, 2)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "talker type", cex = textcex)

cds <- c(21, 6, 29, 8)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "divime_talkertype(marvinator = ...)", cex = textcex, family = "mono")

cds <- c(21, 10, 29, 14)
labs <- "yunitator_old_X.rttm\nyunitator_english_X.rttm"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)

cds <- c(21, 16, 29, 20)
labs <- "yunitate.sh\nyunitate.sh ... english"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)


# vocalization classification -------------------------------------------------------------

cds <- c(31, 4, 39, 2)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "classify vocalizations", cex = textcex)

cds <- c(31, 6, 39, 8)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "divime_classify_vox(marvinator = ...)", cex = textcex, family = "mono")

cds <- c(31, 10, 39, 14)
labs <- "vcm_X.rttm\nvcm_english_X.rttm"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)

cds <- c(31, 16, 39, 20)
labs <- "vcm.sh\nvcm.sh ... english"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)


# word count -------------------------------------------------------------

cds <- c(41, 4, 49, 2)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "word count", cex = textcex)

cds <- c(41, 6, 49, 8)
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = "divime_wordcount(speech_annos = ...)", cex = textcex, family = "mono")

cds <- c(41, 10, 49, 14)
labs <- "WCE_noisemesSad_X.rttm\nWCE_opensmileSad_X.rttm\nWCE_tocomboSad_X.rttm\nWCE_yunitator_english_X.rttm"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)

cds <- c(41, 16, 49, 20)
labs <- "WCE_from_SAD_outputs.sh ... noisemesSad\nWCE_from_SAD_outputs.sh ... opensmileSad\nWCE_from_SAD_outputs.sh ... tocomboSad\nWCE_from_SAD_outputs.sh ... english"
if(r) rect(cds[1], cds[2], cds[3], cds[4])
text(x = mean(cds[c(1, 3)]), y = mean(cds[c(2, 4)]), labels = labs, cex = textcex)
```


Here is a tabular overview.

```{r, results='markup', echo=FALSE}
headers <- c("task", "function", "main argument", "requires", "output template", "comment")

sad1 <- c("**speech detection**", "`divime_sad`", "`module = \"noisemes\"`", " ", "noisemesSad_X", "")
sad2 <- c(" ", " ",                           "`module = \"opensmile\"`", " ", "opensmileSad_X", "")
sad3 <- c(" ", " ",                           "`module = \"tocombo\"`", " ", "tocomboSad_X", "")

diar1 <- c("**talker diarization**", "`divime_diarization`", "`speech_annos = \"noisemes\"`", "noisemesSad_X", "diartk_noisemesSad_X", "identification of individuals speakers")
diar2 <- c(" ", " ",                                     "`speech_annos = \"opensmile\"`", "opensmileSad_X", "diartk_noisemesSad_X", "")
diar3 <- c(" ", " ",                                     "`speech_annos = \"tocombo\"`", "tocomboSad_X", "diartk_tocomboSad_X", "")

yuni1 <- c("**talker type**", "`divime_talkertype`", "`marvinator = FALSE`", " ", "yunitator_old_X", "CHI vs FEM vs MAL")
yuni2 <- c(" ", " ",                             "`marvinator = TRUE`", " ", "yunitator_english_X", "")

vcm1 <- c("**vocalization classification**", "`divime_classify_vox`", "`marvinator = FALSE`", "yunitator_old_X", "vcm_X", "CNS vs NCS vs CRY vs OTH")
vcm2 <- c(" ", " ",                                               "`marvinator = TRUE`", "yunitator_english_X", "vcm_marvinator_X", "")


wce1 <- c("**word count**", "`divime_wordcount`", "`speech_annos = \"noisemes\"`", "noisemesSad_X", "WCE_noisemesSad_X", "")
wce2 <- c(" ", " ",                           "`speech_annos = \"opensmile\"`", "opensmileSad_X", "WCE_tocomboSad_", "")
wce3 <- c(" ", " ",                           "`speech_annos = \"tocombo\"`", "tocomboSad_X", "WCE_tocomboSad_X", "")
wce4 <- c(" ", " ",                           "`speech_annos = \"yunitator_english\"`", "yunitator_english_X", "WCE_yunitator_english_X", "")

xtab <- rbind(sad1, sad2, sad3, diar1, diar2, diar3, yuni1, yuni2, vcm1, vcm2, wce1, wce2, wce3, wce4)
colnames(xtab) <- headers
knitr::kable(xtab,row.names = FALSE)
```



# Set up

## Requirements

What you need as requirement is the `vagrant` software and the `VirtualBox` installed, which provide the environment in which you can set up and run your virtual machine, which in effect gives you access to the DiViMe tools. 

If you are on Windows you might also need `git`. If you are on a Mac chances are you have got `git` already.

For more details see https://divime.readthedocs.io/en/latest/.

**For Windows users**: whenever file paths need to be specified please use single forward slashes, e.g. `c:/data/myaudiofiles`

First off, you will need the following R packages installed:

  - `xml2`
  
  - `tuneR`

  - `parallel` if you want to use this package to set up the DiViMe virtual machine from scratch
  
  - `shiny` and `shinyFiles` if you want to use the shiny interactive user interface


Regarding the DiViMe tools themselves, you will require a fair amount of disk space on the drive where you are going to install the virtual machine (25GB+).

You also need sufficient working memory, at least 8GB. 

The DiViMe documentation also lists at least two processor cores as requirement but allows adapting in case your computer only has one core. The function to set up the VM (described below) will check how many cores you have and adapt the installation process if necessary.

Regardless of whether you are using **Windows or MacOS** you will need `vagrant` and `VirtualBox`. I can't be sure whether the package works with all combinations of versions of these two programs, so if you want to play it safe, you use `vagrant` 2.2.4 and `VirtualBox` 5.2.28, which are the versions that I used while writing this package.

If you use **Windows**, you probably also need `git` (https://git-scm.com/download/win), and I expect that version differences here don't make any difference so take the most recent version.

If you use a **Mac**, you might need to install `Xcode` in order to get access to `git`, but I'm not sure about this. So, try the test first that is mentioned just below and only if that fails, go ahead and install `Xcode`. 

Once you installed everything, you can check whether so far all is set up correctly by using (you probably need to restart your computer though):

```{r}
system2(Sys.which("vagrant"), "-v")
system2(Sys.which("git"), "--version")
```

If this results in some meaningful output, i.e. you get some info about about the versions installed, then you are good to go.

## Set up the DiViMe virtual machine

This step describes the function that sets up the virtual machine. You can also omit this step and follow the instructions described here: https://divime.readthedocs.io/en/latest/install.html. In either case you need a running internet connection. If you already have a virtual machine installed, you can skip this section and go ahead to the next about testing whether your virtual machine can be accessed from R.

The function `divime_vagrant_setup` will take care of these steps without need for using bash commands. There are three things to note here:

  - You need to choose a location where to install the virtual machine. (remember that the drive needs plenty free space)
  
  - By default, the virtual machine gets allocated 2GB of RAM. However, this might be not be enough to run all the DiViMe tools. You can (and probably should) set the memory higher with the argument `memoryoverride`, to about half of the RAM that your computer has. For example, if you have 8GB of RAM, set `memoryoverride = 4096`.
  
  - Finally, note that the process of setting up the virtual machine will take quite some time: expect at least 20 minutes, up to an hour.

```{r, eval = FALSE}
# with a Mac
divime_vagrant_setup(divime_loc = "/Volumes/Data/VMX", memoryoverride = 4096)
# with Windows
divime_vagrant_setup(divime_loc = "C:/Data/VMX", memoryoverride = 4096)
```

## Testing your setup

### Testing the virtual machine

Once the setup step is completed, let's test whether we can communicate with the virtual machine from within R. For this we use `divime_vagrant_state()`, which allows checking or changing the state of the virtual machine. Note that you need to ammend the path of the installation with `DiViMe/`. 

```{r, eval=FALSE}
# with a Mac
divime_vagrant_state(divime_loc = "/Volumes/Data/VMX/DiViMe/", what = "status")
# with Windows
divime_vagrant_state(divime_loc = "C:/Data/VMX/DiViMe/", what = "status")
```

If this results in a printed message "`running (virtualbox)`", things are looking good. If you don't get this message, try:

```{r, eval=FALSE}
# with a Mac
divime_vagrant_state(divime_loc = "/Volumes/Data/VMX/DiViMe/", what = "start")
# with Windows
divime_vagrant_state(divime_loc = "C:/Data/VMX/DiViMe/", what = "start")
```


### Testing the DiViMe tools

If the above steps were successful, you can run the test protocol of the DiViMe package. Running this will take a few minutes because it will download some example data.

```{r, eval=FALSE}
# with a Mac
divime_test(divime_loc = "/Volumes/Data/VMX/DiViMe/")
# with Windows
divime_test(divime_loc = "C:/Data/VMX/DiViMe/")
```

This step downloads a sound file from TalkBank and runs this file (a 5-minute segment) through most of the tools in the DiViMe package. The output of this function should be self-explanatory if successful.

# Using the `divime_` functions

Currently, there are the following functions available to run specific DiViMe tools:

  - `divime_sad()`: runs SAD modules with the option to choose between `opensmile`, `noisemes` and `tocombo`
  
  - `divime_diarization()`: runs `diartk` with your choice of existing SAD output

  - `divime_talkertype()`: runs `yunitate`
  
  - `divime_classify_vox()`: runs `vcm`
  
  - `divime_fullnoisemes()`: runs the `noisemes` module but outputs all found noiseme types (including non-speech)
  
  - `divime_wordcount()`: runs `wce` (word count), which requires speech detection output from either `divime_sad()` or `divime_talkertype(marvinator = TRUE, ...)`
  
  - `divime_run_all()`: runs all the above tools in one go
  
The general idea of how these functions work is the same for all:

  - Specify where your VM is located.

  - Specify a location where you have your audio files stored.
  
  - for `divime_sad()` you need to specify which SAD module you wish to run with the `module=` argument
  
  - for `divime_diarization()` you need to specify what kind of existing SADs you want to use with the `speech_annos=` argument (sorry for the different argument names, I will eventually make them the same for both functions...)
  
What the functions then do is to copy your audio file to your DiViMe location, have the DiViMe tools process them there, and copy the output (.rttm file(s)) back to the location of your audio files. After that is done, the files will be deleted from the DiViMe location again.

During these steps there are again a few things to note:
  
  - one major difference of this package to the instructions of the DiViMe docs is that you can keep your files wherever you want, i.e. you don't need to copy them manually to the DiViMe folder...
  
  - one consequence of this is that you should in fact store your files *outside* the DiViMe location because files in the DiViMe/data folder will be deleted... 
  
  - also note that if you run any of the `divime_` functions, the default is that any existing output is *not* overwritten if it already exists. You can/have to (if desired) force overwriting of existing results (i.e. .rttm files) by setting the `overwrite=` argument to `TRUE`.

In practice, working with the `divime_` functions could look like this.

```{r, eval = FALSE}
# Mac
myaudio <- "/Volumes/Data/audio"      # location for audio files
divime <- "/Volumes/Data/VMX/DiViMe/" # location for DiViMe VM

# Windows
myaudio <- "D:/Data/audio"      # location for audio files
divime <- "C:/Data/VMX/DiViMe/" # location for DiViMe VM
```

If your audio location contains two .wav files, e.g. 'test1.wav' and test2.wav', running the following command will add two more files to this location ('opensmileSad_test1.rttm' and 'opensmileSad_test2.rttm').

```{r, eval = FALSE}
divime_sad(divime_loc = divime, audio_loc = myaudio, module = "opensmile")
```

If you then want to run the diarization tool, the following function will add two more files to your audio location ('diartk_opensmileSad_test1.rttm' and 'diartk_opensmileSad_test2.rttm').

```{r, eval=FALSE}
divime_diarization(divime_loc = divime, audio_loc = myaudio, speech_annos = "opensmile")
```


Let's go for a *proper* example, i.e. using two of the example files provided in the package. These short files contain a few sentences spoken by a voice synthesizer. We start by setting up temporary folder and copy the sound files there. (Note that I use the R temporary file folder for this, which is deleted once you close R).


```{r, results='hide'}
tdir <- file.path(normalizePath(tempdir(), winslash = "/"), "avutils")
dir.create(tdir)
file.copy(from = system.file("synthetic_speech.wav", package = "avutils"),
          to = file.path(tdir, "synthetic_speech.wav"))
file.copy(from = system.file("synthetic_speech_overlap.wav", package = "avutils"),
          to = file.path(tdir, "synthetic_speech_overlap.wav"))
```

```{r, eval= FALSE}
# open location in file browser
shell.exec(tdir) # Windows
system2("open", tdir) # Mac
```

```{r, echo=FALSE, results='hide'}
file.copy(from = system.file("opensmileSad_synthetic_speech.rttm", package = "avutils"),
          to = file.path(tdir, "opensmileSad_synthetic_speech.rttm"))
file.copy(from = system.file("opensmileSad_synthetic_speech_overlap.rttm", package = "avutils"),
          to = file.path(tdir, "opensmileSad_synthetic_speech_overlap.rttm"))
file.copy(from = system.file("noisemesSad_synthetic_speech.rttm", package = "avutils"),
          to = file.path(tdir, "noisemesSad_synthetic_speech.rttm"))
file.copy(from = system.file("noisemesSad_synthetic_speech_overlap.rttm", package = "avutils"),
          to = file.path(tdir, "noisemesSad_synthetic_speech_overlap.rttm"))
```


Now let's run two speech detection algorithms on the two files in this folder.

```{r, eval=FALSE}
divime_sad(audio_loc = tdir, divime_loc = "/Volumes/Data/VMX/DiViMe/", module = "opensmile", vmshutdown = FALSE)
divime_sad(audio_loc = tdir, divime_loc = "/Volumes/Data/VMX/DiViMe/", module = "noisemes", vmshutdown = FALSE)
```

This should lead to four `.rttm` files being created in the data folder:

```{r}
list.files(tdir)
```

Now let's read back one of these files.

```{r}
file1 <- read_rttm(file.path(tdir, "opensmileSad_synthetic_speech.rttm"))
file1
```

Please note that during the reading with `read_rttm()` I assign column names that are not present in the actual .rttm file and I create an extra column to set the end time of each segment.

```{r, echo=FALSE}
cat(paste(readLines(file.path(tdir, "opensmileSad_synthetic_speech.rttm")), collapse = "\n"))
```










# DiViMe shiny app

The app is still in an early development stage.


```{r, eval=FALSE}
divime_app() # starts app in your default browser
divime_app(launchinbrowser = FALSE) # starts app in RStudio window (if you use RStudio)
```




# Evaluation of speech detection (NEEDS MAJOR REWORKING!!!)

We start by providing the paths to two annotation files (one manually coded with ELAN, one the output from noisemes), which are part of the package and relate to an audio file that I made up with synthetic speech.

```{r}
# a 'test' file
x1 <- system.file("noisemesSad_synthetic_speech.rttm", package = "avutils")
# the gold standard reference (manual coding)
x2 <- system.file("synthetic_speech.eaf", package = "avutils")
```

The first example is 'normal' in that both files contain detected speech and the overlap between the two annotation sets is fairly good.

```{r, out.width="50%"}
evaluate_sad(test = x1, reference = x2, overlap_res = 500)
```

Let's first consider the graphical output. We see two sets of lines, one black and one gold, which refer to the 'test' file (`x1`) and the reference file (or the 'gold standard', `x2` in our case). The lines represent the speech detected in both annotation sets. When you use ELAN files, note that the function will take all annotations from all tiers and consider them speech. If you want to exclude certain annotations or tiers, you can exclude them by providing the relevant arguments (see `?collapse_tiers` for more details on this). On the left side, we can see the time stamps (in seconds) that are covered. 

The next thing to note is the top horizontal axis (evaluation points). The corresponding argument (`overlap_res=`) controls how many evaluation points are used. At each evaluation point, the function compares the two annotation sets with respect to whether an annotation exists at this point or not. This is somewhat contrasting the evaluation pipeline in the Python version, where each individual sample is evaluated.

The last type of information provided by the graphic are red circles between the two line sets. These circles (if present) indicate a mismatch between the two annotation sets, i.e. one file has speech detected at this point while the other has not.

The numerical output provides three types of information. First, the *false alarm* rate, which takes all evaluation points that show no speech in the standard `x2` and calculates the proportion of evaluation points that (falsly) contain speech in the test file `x1`. The *missed speech* rate is the opposite, that is the proportion of no-speech points in `x1` when there is speech in the standard `x2`. Finally, there is the *total* congruence, which is just the total proportion of all evaluation points where `x1` has the same value as `x2`. So, for FA and MSR, small values close to 0 are better, while for the total 1 is the best value. 

For a proper evaluation, keep in mind to set `overlap_res=` to a reasonably large value. One rule of thumb could be to use ten evaluation points per second recording, which would provide precision to the tenth of a second. 

## Visual versus numerical results

In order to aid visual assessment, `evaluate_sad()` allows to break the display into multiple rows and/or to select a time segment within the file. Note that when you 'break' the file over multiple segments, all the evaluation metrics are returned for each segment separately. In the specific example that follows, note that in the last segment, there are no false alarms at all.

```{r}
# break the display over 5 segments
evaluate_sad(test = x1, reference = x2, nsegments = 5, overlap_res = 100)
```

You can also select a time window. (the example here corresponds roughly to the second segment from just above).

```{r}
evaluate_sad(test = x1, reference = x2, from = 9.46, to = 18.94, overlap_res = 100)
```


## Special cases

The DiViMe manual considers special cases when one or two of the files submitted contain no speech. Let's look at this with an example where we use predefined segments of the example files.

First, we consider the special case where the gold/manual annotation is empty and also the SAD module detected no speech.

```{r, out.width="50%"}
evaluate_sad(test = x1, reference = x2, nsegments = 1, from = 27, to = 28, overlap_res = 10)
```

As expected, both FA and MSR are 0 (as defined in the DiViMe manual).

Second, we consider the next case where the gold/manual annotation is empty but the SAD module detected some speech.

```{r, out.width="50%"}
evaluate_sad(test = x1, reference = x2, nsegments = 1, from = 27, to = 28.3, overlap_res = 10)
```

Here we see that the false alarm rate is 1 while the missed speech rate is set to 0.

And finally, the case where there was speech in the gold standard, but the SAD module didn't detect anything:

```{r, out.width="50%"}
evaluate_sad(test = x1, reference = x2, nsegments = 1, from = 23.2, to = 25, overlap_res = 10)
```

Here, FA = 0 and MSR is 1.

## splitting files

```{r, echo=2:4, fig.width=8, fig.height=3.5, out.width="100%"}
par(mfrow = c(1, 3), family = "serif")
evaluate_sad(test = x1, reference = x2, nsegments = 2, from = 27, to = 28, overlap_res = 10)
evaluate_sad(test = x1, reference = x2, nsegments = 2, from = 27, to = 28.3, overlap_res = 10)
evaluate_sad(test = x1, reference = x2, nsegments = 2, from = 23.2, to = 25, overlap_res = 10)
```



